*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-21 Code/Manuscript_O2.Rmd
edits to PUR materials & methods
editing growth graph results

PurParPlot_cap <- glue("**The absorbed Photosynthetically Usable Radiation (PUR) (µmol photons m^-2^ s^-1^) vs. the Photosynthetically Active Radiation (PAR) (µmol photons m^-2^ s^-1^).**",
                       "The correlation between PAR, plotted on the x-axis and absorbed PUR, plotted on the y-axis, are colored for each growth spectral waveband; blue for 450 nm, red for 660 nm and black for white LED full spectrum light.",
                       "The grey dashed line represents a hypothetical one to one correlation.",
                       "**A.** is *Prochlorococcus marinus*  MED4.",
                       "**B.** is *Prochlorococcus marinus*  SS120.",
                       "**C.** is *Prochlorococcus marinus*  MIT9313.",

@@

%%%% OceanProteinPortal data show complexes in environment  %%%

Furthermore,
Evidence of proteins
derived from HL ecotypes of *P. marinus* in OMZ at depths up to 200
meters,  with O~2~ of 15 µM. The extent to which [O~2~] defines
the niches occupied by different *P. marinus* ecotypes, as compared to
potentially covarying environmental variables like photoperiod, light
spectrum, and light level, is poorly described.

@@

ProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4, SS120 and MIT9313.",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "Culture growth experimental conditions indicated by horizontal grey lines for depths approximating Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^) and vertical grey lines for O~2~ (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

### MED4 PAR and growth trends {.unnumbered}

@@

The growth rate of MED4, from the High-Light (HLI) clade of *P.
marinus*, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), when grown under 250 µM O~2~. No growth was
observed under any imposed conditions under a 4 h photoperiod. MED4 was
able to grow under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ but only when the blue light photoperiod was longer than 4 h; the
white LED light photoperiod was longer than 8 h; or when the red light
photoperiod was longer than 16 h. Growth rate increased
with increasing photoperiods for all spectral wavebands when exposed to
PAR of 90 and 180 µmol photons m^-2^ s^-1^. MED4 grew fastest for each
spectral waveband under 16 h photoperiods. The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue
light and 16 h photoperiod.

@@

Similar to the growth trends of the 250 µM O~2~ experiments, MED4
maintained at 25 µM O~2~, showed the fastest growth rate when the
photoperiod was 16 h for each spectral waveband, across PAR levels
(Figure \@ref(fig:MED4PAR)). The 25 µM O~2~ growth experiments showed
generally similar growth rate as the 250 µM O~2~ growth experiments,
with the interesting exception that growth rates increased by 40% under
25 µM O~2~ at PAR of 30 µmol photons m^-2^ s^-1^ and 16 h photoperiod,
and even showed growth under 12 h red light photoperiod, in contrast to
the 250 µM O~2~ growth experiments. White LED light treatments grew
slower, if at all, under 25 µM O~2~, only growing under higher PAR of 90
and 180 µmol photons m^-2^ s^-1^ and longer photoperiods of 12 and 16 h.
As under 250 µM O~2~, the fastest growth rate observed for each spectral
waveband at each imposed PAR were greatest under the 16 h photoperiod.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 180 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod
experiments under white LED light were not performed because no growth
was achieved when grown under an 8 h photoperiod.

@@

MED4 did not grow under any tested conditions sparged to the lowest O~2~
of 2.5 µM (Figure \@ref(fig:MED4PAR)). 2.5 µM O~2~ growth experiments
were not conducted for 4 and 16 h photoperiods, as no growth occurred
when MED4 was exposed to 8 and 12 h photoperiods.

@@

### SS120 PAR and growth trends {.unnumbered}

@@

The growth rate of SS120, from the Low-Light (LLIII) clade of *P.
marinus*, increased with lower imposed PAR and longer photoperiods, when
grown under 250 µM O~2~ (Figure \@ref(fig:SS120PAR)). SS120 was able to
grow under the lowest PAR treatment of 30 µmol photons m^-2^ s^-1^ but
only when the photoperiod was longer than 4 h under blue, white, and red
light. SS120 required longer than 8 h photoperiod of imposed green
light. No growth was observed under any blue light photoperiods when
exposed to PAR of 90 µmol photons m^-2^ s^-1^. Growth rate increased
with increasing photoperiods for white, red and green light when exposed
to PAR of 90 µmol photons m^-2^ s^-1^; however, no growth was observed
when the red light photoperiod was 16 h. Experiments conducted under PAR
of 180 µmol photons m^-2^ s^-1^ showed no growth under imposed blue and
red lights. Growth rate decreased with longer photoperiods and showed
growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol
photons m^-2^ s^-1^ and white LED light. The growth rate generally
increased with longer photoperiods under green light treatments, however
no growth was observed under the 8 h photoperiod. The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 90 µmol photons
m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

Similar to the growth trends of the 250 µM O~2~ and PAR of 30 µmol
photons m^-2^ s^-1^ experiments, SS120 maintained at 25 µM O~2~, no
growth was observed under a 4 h photoperiod for imposed red spectral
waveband, however under blue light, SS120 was able to grow (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM
O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew
under 4 and 8 h blue light and 16 h red light photoperiods. The growth
rate decreased with increasing photoperiod under PAR of 90 µmol photons
m^-2^ s^-1^ white LED light treatments. Blue light treatments under PAR
of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h
photoperiod. This growth rate decreased by 53% compared to the 8 h blue
light photoperiod under PAR of 90 µmol photons m^-2^ s^-1^. The µ~max~
was
`r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod.

The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^ experiments under white LED
light were not performed due to time constraints.

@@

SS120 only grew
(`r PURPARFits %>% filter(Strain == "SS120", WL == "660", O2_uM == "2.5", Photoperiod == 12) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^) under PAR of 30 µmol photons m^-2^ s^-1^ and 12 h red light
photoperiod when sparged to the lowest O~2~ of 2.5 µM (Figure
\@ref(fig:SS120PAR)). SS120 did not grow under any other tested
conditions. 2.5 µM O~2~ growth experiments were not conducted for 4 and
16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^, as no
growth occurred when SS120 was exposed to 8 and 12 h photoperiods. Red
light 16 h photoperiod experiments were not performed due to time
constraints.

@@

\newpage
### MIT9313 PAR results {.unnumbered}

@@

The growth rate of MIT9313, from the Low-Light (LLIV) clade of *P.
marinus*, increased with low to moderate imposed PAR and longer
photoperiods, when grown under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)). Growth rate increased with increasing
photoperiods under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ for blue, white LED and red light. The green light treatment
showed growth only when the photoperiod was longer than 12 h and
decreased by \~ 50% under a 16 h photoperiod. Under PAR of 90 µmol
photons m^-2^ s^-1^, growth rate increased with longer photoperiods for
white LED, red, and green light treatments. Growth rates decreased from
4 to 8 h blue light photoperiod and exhibited no growth under
photoperiods of 12 and 16 h. Blue light did not induce growth at 180
µmol photons m^-2^ s^-1^ under any photoperiod; white LED and red light
treatments only exhibited growth under the 8 h photoperiod and growth
rates increased with longer photoperiods under green light. The µ~max~
was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 30 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod.

>>>>

PurParPlot_cap <- glue("**Absorbed peak Photosynthetically Usable Radiation (PUR) (µmol photons m^-2^ s^-1^) vs. peak Photosynthetically Active Radiation (PAR) (µmol photons m^-2^ s^-1^).**",
                       "The correlation between PAR, plotted on the x-axis and absorbed PUR, plotted on the y-axis, are colored for each growth spectral waveband; blue for 450 nm, red for 660 nm and black for white LED full spectrum light.",
                       "The grey dashed line represents a hypothetical one to one correlation.",
                       "**A.** is *Prochlorococcus marinus*  MED4.",
                       "**B.** is *Prochlorococcus marinus*  SS120.",
                       "**C.** is *Prochlorococcus marinus*  MIT9313.",

@@

## Detection of *Prochlorococcus* proteins across O~2~ and light niches in the ocean

Proteins derived from Clade HLXXX ecotypes of *P. marinus* were detected in OMZ at depths up to 200 meters,  with O~2~ of 15 µM. The extent to which [O~2~] defines
the niches occupied by different *P. marinus* ecotypes, as compared to
potentially covarying environmental variables like photoperiod, light
spectrum, and light level, is poorly described.

@@

ProchloroPhotosynthDepthO2_cap <- glue("**Ocean detection of *Prochlorococcus marinus* photosynthesis complexes.**", 
                       "Protein detections are plotted vs. O~2~ (µM) (X axis) and depth (m) (Y axis) at sample origin.",
                       "Rows separate data annotated as from *Prochlorococcus marinus* strains MED4, SS120 and MIT9313.",
                       "Columns show detections of proteins annotated as  Photosystem II (PSII), Cytochromeb6f complex (Cytb6f), Photosystem I (PSI), ATP Synthase or Ribulose-1,5-bisphosphate oxygenase carboxylase (RUBISCO).",
                       "For comparison culture growth experimental conditions are indicated by horizontal grey lines for depths approximating imposed peak Photosynthetically Active Radiation (µmol photons m^-2^ s^-1^); and vertical grey lines for O~2~ (µM).",
                       "Point colours identify *Prochlorococcus* clade assignments",
                       "Data from OceanProteinPortal (https://www.oceanproteinportal.org/).",
 .sep = " ")

@@

## *Prochlorococcus marinus* growth responses to photoperiod, PAR and O~2~

### MED4 Clade HLI {.unnumbered}

@@

The growth rates of *P. marinus* MED4, Clade HLI, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), when grown under 250 µM O~2~. No growth was
observed under any imposed conditions under a 4 h photoperiod. MED4 was
able to grow under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ but only when the blue light photoperiod was longer than 4 h; the
white LED light photoperiod was longer than 8 h; or when the red light
photoperiod was longer than 16 h. Growth rate increased
with increasing photoperiods for all spectral wavebands when exposed to
PAR of 90 and 180 µmol photons m^-2^ s^-1^. MED4 grew fastest for each
spectral waveband under 16 h photoperiods. The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue
light and 16 h photoperiod.

@@

Similar to the growth trends under 250 µM O~2~, MED4
maintained at 25 µM O~2~ showed the fastest growth rates when the
photoperiod was 16 h for each spectral waveband, across PAR levels
(Figure \@ref(fig:MED4PAR)). The 25 µM O~2~ growth experiments showed
generally similar growth rates as the 250 µM O~2~ growth experiments,
with the interesting exception that growth rates increased by 40% under
25 µM O~2~ at PAR of 30 µmol photons m^-2^ s^-1^ and 16 h photoperiod,
and even showed growth under 12 h red light photoperiod, in contrast to
the 250 µM O~2~ growth experiments. White LED light treatments grew
slower, if at all, under 25 µM O~2~, only growing under higher PAR of 90
and 180 µmol photons m^-2^ s^-1^ and longer photoperiods of 12 and 16 h.
As under 250 µM O~2~, the fastest growth rate observed for each spectral
waveband at each imposed PAR were greatest under the 16 h photoperiod.
The µ~max~ was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 180 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod
experiments under white LED light were not performed because no growth
was achieved when grown under an 8 h photoperiod.

@@

MED4 did not grow under when sparged to the lowest O~2~
of 2.5 µM (Figure \@ref(fig:MED4PAR)). 2.5 µM O~2~ growth experiments
were not conducted for 4 and 16 h photoperiods, as no reproducible growth occurred
when MED4 was exposed to 8 and 12 h photoperiods.

@@

### SS120 Clade LLIII {.unnumbered}

@@

The growth rates of *P. marinus* SS120, Clade LLIII, increased with lower imposed PAR and with longer photoperiods, when
grown under 250 µM O~2~ (Figure \@ref(fig:SS120PAR)). SS120 was able to
grow under the lowest PAR treatment of 30 µmol photons m^-2^ s^-1^ but
only when the photoperiod was longer than 4 h under white and red
light. No growth was observed under any blue light photoperiods when
exposed to PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate increased
with increasing photoperiods for white and red light when exposed
to PAR of 90 µmol photons m^-2^ s^-1^; however, no growth was observed
when the red light photoperiod was 16 h. Experiments conducted under PAR
of 180 µmol photons m^-2^ s^-1^ showed no growth under blue nor
red lights. Growth rate decreased with longer photoperiods and showed
growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol
photons m^-2^ s^-1^  white LED light.  The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 90 µmol photons
m^-2^ s^-1^ white LED light and 16 h photoperiod.

@@

Similar to the growth trends under the 250 µM O~2~ and PAR of 30 µmol
photons m^-2^ s^-1^ experiments, SS120 maintained at 25 µM O~2~ showed no
growthunder a 4 h photoperiod for imposed red spectral
waveband, however under blue light, SS120 was able to grow (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM
O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew
under 4 and 8 h blue light and 16 h red light photoperiods. The growth
rate decreased with increasing photoperiod under PAR of 90 µmol photons
m^-2^ s^-1^ white LED light treatments. Blue light treatments under PAR
of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h
photoperiod. The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod.

The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^  under white LED light experiments were not performed due to time constraints.

@@

SS120 did not reproducibly grow when sparged to the lowest O~2~ of 2.5 µM (Figure
\@ref(fig:SS120PAR)). 2.5 µM O~2~ growth experiments were not conducted for 4 and
16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^, as no
growth occurred when SS120 was exposed to 8 and 12 h photoperiods. Red
light 16 h photoperiod experiments were not performed due to time constraints.

@@

### MIT9313 Clade LLIV {.unnumbered}

@@

The growth rates of *P. marinus* MIT9313, Clade LLIV, increased with low to moderate imposed PAR and longer
photoperiods, when grown under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)). Growth rate increased with increasing
photoperiods under the lowest PAR treatment of 30 µmol photons m^-2^
s^-1^ for blue, white LED and red light. The green light treatment
showed growth only when the photoperiod was longer than 12 h and
decreased by \~ 50% under a 16 h photoperiod. Under PAR of 90 µmol
photons m^-2^ s^-1^, growth rate increased with longer photoperiods for
white LED, red, and green light treatments. Growth rates decreased from
4 to 8 h blue light photoperiod and exhibited no growth under
photoperiods of 12 and 16 h. Blue light did not induce growth at 180
µmol photons m^-2^ s^-1^ under any photoperiod; white LED and red light
treatments only exhibited growth under the 8 h photoperiod and growth
rates increased with longer photoperiods under green light. The µ~max~
was
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 30 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod.
